An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-07-15 EconGrowth.Rnw
Improved description of CES production function.

\section*{To-do List}
\begin{itemize}
\item{Heun: Look into constraining $\delta \in [0,1]$
and $\delta_1 \in [0,1]$ in the CES and CES$e$ models.}
\item{Pruim: Respond to Heun's wish list for graphs.}
\item{Heun: Figure out why CESe for ZA with $x$ fails during resampling.}
\item{Heun: Review all assumptions for CES with energy models. 
There are several assumptions for rho and rho1 for various countries. 
When we are resampling, we probably don't want to specify those values.}
\item{Heun: Investigate possible conflicting use of exergy by Warr and Ayres.}
\item{Heun: Fix the energy sankey diagram from the seminar}
\item{Heun: Include adjusted sankey diagram in the seminar}
\item{Heun: Add CES production function fits}
      \begin{itemize}
      \item{Write up the decisions about parameter estimation for the 
            CES production function fits}
      \item{Revisit my assumptions for the CES production function 
            (with energy) for each country.}
      \item{Fix the problems with $\sigma_1$ and $\sigma$ values in 
            the CES parameter tables in the appendix}
      \item{Add CES rows to the AIC table. Code is in the 
            \texttt{createAICTable} function, but it is not yet working. 
            Need to find a way to calculate AIC for the object that is 
            returned from the \texttt{cesEst} function.}
      \end{itemize}
\item{Heun: Fix warnings of the form \texttt{Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562} in the code that generates the Cobb-Douglas with energy fits.}
\item{de Wit: review, critique, and contextualize.}
\item{\st{Heun: Add option for $m = 1$ in single-factor models.}}
\item{\st{Heun: Re-implement confidence interval calculations to use
resampling methods.}}
      \begin{itemize}
      \item{\st{SF}}
      \item{\st{CD}}
      \item{\st{CDe}}
      \item{\st{CES}}
      \item{\st{CESe}}
      \item{\st{LINEX}}
      \end{itemize}
\item{\st{Heun: Include scatter graphs for model parameters 
estimated by resampling methods.}}
      \begin{itemize}
      \item{\st{SF}}
      \item{\st{CD}}
      \item{\st{CDe}}
      \item{\st{CES}}
      \item{\st{CESe}}
      \item{\st{LINEX}}
      \end{itemize}
\item{\st{Heun: Decide if the single factor models should include constraints
on $\alpha$, $\beta$, and $\gamma$. At present, the single-factor model allows
values outside the ranges $\alpha \in (0,1)$, $\beta \in (0,1)$, and $\gamma \in (0,1)$.}}
\item{\st{Heun: Make smaller scatter plots for resample coefficients for those 
countries where $u$ is involved.}}
\item{\st{Heun: Add a table that shows the fraction of unconverged resample fits
for each model, energy type, and factor.}}
\item{\st{Heun: Get AIC calculations working again.}}
\item{\st{Pruim: Add base fit red dots to all cloud plots.}}
      \begin{itemize}
      \item{\st{SF}}
      \item{\st{CD}}
      \item{\st{CDe}}
      \item{\st{CES}}
      \item{\st{CESe}}
      \item{\st{LINEX}}
      \end{itemize}
\end{itemize}

@@

Constraints on the fitting parameters include 
$\delta_1 \in (0,1)$,
$\delta \in (0,1)$,
$\rho_1 \in (-1,0) \cup (0,\infty)$, and
$\rho \in (-1,0) \cup (0,\infty)$.
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, 
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$, 
$\delta_1 \in [0, 1]$ describes the relative importance of capital ($k$)
and labor ($l$), 
$\delta \in [0, 1]$ describes the importance 
of ($kl$) relative to ($e$), 
and $\gamma$ is a fitting parameter that represents productivity. 
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.

>>>>

\section*{To-do List}
\begin{itemize}
\item{Heun: Look into constraining $\delta \in [0,1]$
and $\delta_1 \in [0,1]$ in the CES and CES$e$ models.}
\item{Heun: Investigate why wew obtain $\rho_1 < -1$ 
for some US CES$e$ resample models.}
\item{Pruim: Respond to Heun's wish list for graphs.}
\item{Heun: Figure out why CESe for ZA with $x$ fails during resampling.}
\item{Heun: Review all assumptions for CES with energy models. 
There are several assumptions for rho and rho1 for various countries. 
When we are resampling, we probably don't want to specify those values.}
\item{Heun: Investigate possible conflicting use of exergy by Warr and Ayres.}
\item{Heun: Fix the energy sankey diagram from the seminar}
\item{Heun: Include adjusted sankey diagram in the seminar}
\item{Heun: Add CES production function fits}
      \begin{itemize}
      \item{Write up the decisions about parameter estimation for the 
            CES production function fits}
      \item{Revisit my assumptions for the CES production function 
            (with energy) for each country.}
      \item{Fix the problems with $\sigma_1$ and $\sigma$ values in 
            the CES parameter tables in the appendix}
      \item{Add CES rows to the AIC table. Code is in the 
            \texttt{createAICTable} function, but it is not yet working. 
            Need to find a way to calculate AIC for the object that is 
            returned from the \texttt{cesEst} function.}
      \end{itemize}
\item{Heun: Fix warnings of the form \texttt{Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562} in the code that generates the Cobb-Douglas with energy fits.}
\item{de Wit: review, critique, and contextualize.}
\item{\st{Heun: Add option for $m = 1$ in single-factor models.}}
\item{\st{Heun: Re-implement confidence interval calculations to use
resampling methods.}}
      \begin{itemize}
      \item{\st{SF}}
      \item{\st{CD}}
      \item{\st{CDe}}
      \item{\st{CES}}
      \item{\st{CESe}}
      \item{\st{LINEX}}
      \end{itemize}
\item{\st{Heun: Include scatter graphs for model parameters 
estimated by resampling methods.}}
      \begin{itemize}
      \item{\st{SF}}
      \item{\st{CD}}
      \item{\st{CDe}}
      \item{\st{CES}}
      \item{\st{CESe}}
      \item{\st{LINEX}}
      \end{itemize}
\item{\st{Heun: Decide if the single factor models should include constraints
on $\alpha$, $\beta$, and $\gamma$. At present, the single-factor model allows
values outside the ranges $\alpha \in (0,1)$, $\beta \in (0,1)$, and $\gamma \in (0,1)$.}}
\item{\st{Heun: Make smaller scatter plots for resample coefficients for those 
countries where $u$ is involved.}}
\item{\st{Heun: Add a table that shows the fraction of unconverged resample fits
for each model, energy type, and factor.}}
\item{\st{Heun: Get AIC calculations working again.}}
\item{\st{Pruim: Add base fit red dots to all cloud plots.}}
      \begin{itemize}
      \item{\st{SF}}
      \item{\st{CD}}
      \item{\st{CDe}}
      \item{\st{CES}}
      \item{\st{CESe}}
      \item{\st{LINEX}}
      \end{itemize}
\end{itemize}

@@

$\gamma$ is a fitting parameter that represents productivity. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, and
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$. 
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
Similarly, as $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
$\delta_1$ describes the relative importance of capital ($k$)
and labor ($l$), and
$\delta$ describes the importance of ($kl$) relative to ($e$).

Constraints on the fitting parameters include 
$\delta_1 \in (0,1)$,
$\delta \in (0,1)$,
$\rho_1 \in (-1,0) \cup (0,\infty)$, and
$\rho \in (-1,0) \cup (0,\infty)$. **** Describe here how 
we implemented the constraints. ****
